The \dealii{} Library, Version 9.3

### 2021-06-20 9.3/paper.tex
description of FEInterfaceValues

\author[9,12]{Peter Munch}
 \affil[12]{Institute of Material Systems Modeling,
 Helmholtz-Zentrum Hereon,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hereon.de}}}

@@

The major changes of this release are:
%
\begin{itemize}
  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%
In addition, we discuss the \candi{} installation program in Section~\ref{subsec:candi}.

@@

Additionally, the particle infrastructure now supports a faster search algorithm
for particles that moved farther than one cell width between particle sorting
operations, and support for updating ghost particles (particles that live in
ghost cells around the local domain) by updating their properties instead of
destroying and rebuilding their container. The latter step improves the
efficiency of ghost particle exchange significantly. With this feature,
\dealii can be used to generate scalable parallel code for Lagrangian
models such as the Discrete Element Method (DEM) \cite{golshan2021lethedem}
or Molecular Dynamics.

@@

To give an example of the new features, let us consider the evaluation of a
surface tension force in the context of sharp-interface methods, whose
contribution is added to a fluid solver by multiplication with test function
and addition over quadrature points located at the interface section
$\Gamma_K=\Gamma \cup K$ located on the current cell $K$ and that is,
in general, located differently within $K$ than for any other cell:
\begin{align*}
\left(\mathbf{v}, \kappa \, \mathbf{n}\right)_\Gamma
\approx
\sum_q \mathbf{v}(\mathbf{x}_q) \cdot \left(\kappa(\mathbf{x}_q) \, \mathbf{n}(\mathbf{x}_q)\right) (JxW)_q.
\end{align*}
In \dealii{}, this can now be conveniently written as
\begin{c++}
phi_curvature.reinit(cell, reference_points);
phi_normal.reinit(cell, reference_points);
phi_force.reinit(cell, reference_points);

phi_curvature.evaluate(curvature_values, EvaluationFlags::values);
phi_normal.evaluate(normal_values, EvaluationFlags::values);

for (unsigned int q = 0; q < n_points; ++q)
  phi_force.submit_value(phi_curvature.get_value(q) *
                         phi_normal.get_value(q) * JxW[q], q);

phi_force.integrate(force_values, EvaluationFlags::values);
\end{c++}
The quadrature points (at reference positions
\texttt{reference\_points}) and the related \texttt{JxW} value can,
for example, come from
a mesh of lower dimension. Determining to which \texttt{cell} a quadrature
point belongs to on the background mesh, including the reference cell
coordinates \texttt{reference\_\allowbreak points}, is aided by functions like
\texttt{find\_active\_cell\_around\_point()}
\texttt{find\_all\_active\_cells\_\allowbreak around\_\allowbreak point()} from the \texttt{GridTools} namespace. While these
functions have been available in \dealii{} previously, their performance has
been considerably enhanced with the aforementioned more optimized code paths
for selected mappings.

@@

@@

\item \texttt{step-74} SIPG for the Poisson equation \todo[inline]{TODO}

@@

B.~Blais was partially supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant

>>>>

\author[9,12]{Peter Munch}
 \affil[12]{Institute of Materials Research, Materials Mechanics,
 Helmholtz-Zentrum Geesthacht,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hzg.de}}}

@@

The major changes of this release are:
%
\begin{itemize}
  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
  \item Simplified implementation for cell face integrals (see Section~\ref{subsec:feinterfacevalues});
  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%
In addition, we discuss the \candi{} installation program in Section~\ref{subsec:candi}.

@@

Additionally, the particle infrastructure now supports a faster search algorithm
for particles that moved farther than one cell width between particle sorting
operations, and support for updating ghost particles (particles that live in
ghost cells around the local domain) by updating their properties instead of
destroying and rebuilding their container. The latter step improves the
efficiency of ghost particle exchange significantly.

@@

To give an example of the new features, let us consider the evaluation of a
surface tension force in the context of sharp-interface methods, whose
contribution is added to a fluid solver by multiplication with test function
and addition over quadrature points located at the interface section
$\Gamma_K=\Gamma \cup K$ located on the current cell $K$ and that is,
in general, located differently within $K$ than for any other cell:
\begin{align*}
\left(\mathbf{v}, \kappa \, \mathbf{n}\right)_\Gamma
\approx
\sum_q \mathbf{v}(\mathbf{x}_q) \cdot \left(\kappa(\mathbf{x}_q) \, \mathbf{n}(\mathbf{x}_q)\right) (JxW)_q.
\end{align*}
In \dealii{}, this can now be conveniently written as
\begin{c++}
phi_normal.reinit(cell, reference_points);
phi_force.reinit(cell, reference_points);
phi_normal.evaluate(normal_values, EvaluationFlags::values);
for (unsigned int q = 0; q < n_points; ++q)
   phi_force.submit_value(phi_curvature.get_value(q) *
                          phi_normal.get_value(q) * JxW[q], q);
phi_force.integrate(force_values, EvaluationFlags::values);
\end{c++}
The quadrature points (at reference positions
\texttt{reference\_points}) and the related \texttt{JxW} value can,
for example, come from
a mesh of lower dimension. Determining to which \texttt{cell} a quadrature
point belongs to on the background mesh, including the reference cell
coordinates \texttt{reference\_\allowbreak points}, is aided by functions like
\texttt{find\_active\_cell\_around\_point()}
\texttt{find\_all\_active\_cells\_\allowbreak around\_\allowbreak point()} from the \texttt{GridTools} namespace. While these
functions have been available in \dealii{} previously, their performance has
been considerably enhanced with the aforementioned more optimized code paths
for selected mappings.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Simplied implementation for cell face integrals}
\label{subsec:feinterfacevalues}
Discontinuous Galerkin (DG) methods require evaluations of averages and jumps accross cell faces, therefore we need to compute values of the shape functions and solutions from two adjacent cells. The \texttt{FEInterfaceValues} class is designed to provide an interface to directly obtain jumps and averages. When used with \texttt{MeshWorker::mesh\_loop()}, we can assemble cell face integrals in a uniform way independent of mesh types, which significantly simplifies the implementation.

Internally, this class provides an abstraction for two \texttt{FEFaceValues} objects (or \texttt{FESubfaceValues} when using adaptive refinement). The class introduces new interface degrees of freedom indices that are the union of the degrees of freedom indices of the two \texttt{FEFaceValues} objects. The interface degrees of freedom indices can be converted to the corresponding local degrees of freedom indices of the two cells using a helper function. Furthermore, this class supports vector-valued problems in a similar way as \texttt{FEFaceValues} does. We can easily obtain the corresponding scalar or vector values by providing an \texttt{FEValuesExtractors} object. Overall, this class is intended to be a low level replacement for \texttt{MeshWorker} and \texttt{LocalIntegrators}, and a higher level abstraction compared to assembling face terms manually.

Examples of \texttt{FEInterfaceValues} for scalar problems can be found in \texttt{step-12}, \texttt{step-47}, \texttt{step-50}, and the new tutorial \texttt{step-74}. While examples of vector-valued problems will be added in the future.

@@

\item \texttt{step-74} implements the symmetric interior penalty Galerkin 
(SIPG) method for Poisson's equation using the \texttt{FEInterfaceValues} class
in combination with \texttt{MeshWorker::mesh\_loop()}. This tutorial demonstrates a
simple and mesh-independent way to assemble cell face integrals.

@@
